### 2021-06-11 papers/thesis/full/ch3-disk-slices/disk-slices.tex
Addressed the new minor comments, though I have questions for the other points. Also changed the rec coeff mats defs for each chapter to make consistent.
Edits to Zernike stuff and other minor edits

For the disk, a review of spectral methods on the disk for solving the Poisson equation has been completed \cite{boyd2011comparing}, that looks at methods involving various bases for expanding functions in the disk including Zernike polynomials and Chebyshev polynomials. It is concluded that using Chebyshev polynomials yields mixed results, but that there is also no single advised basis to use. Zernike polynomials were found to yield accurate results, and ideas for how to implement a radial transform have progressed in recent years \cite{slevinsky2018use, sakai2009application, wilber2017computing} \bstodo{not sure what to say here?}. One solution proposed is pairing the Zernike disk OPs with a radial basis defined in terms of Jacobi polynomials \cite{vasil2016tensor}:
\bseqnnumber{
	Q_n^{k,m}(r) := C_n^{k,m} r^m P^{(k,m)}_n(2r^2 - 1), \label{eqn:ds:diskops}
}
where $C_n^{k,m}$ is some normalising constant so that they are orthonormal with respect to $\int_0^1 Q_n^{k,m}(r) \; Q_{n'}^{k,m}(r) \; (1-r^2)^k \; r \; \D r = \delta_{n,n'}$. $\{Q_n^{k,m}\}$ are therefore a two-parameter family of orthogonal polynomials. The use of polar coordinates here then allows for radially symmetric partial differential (and other) operators, as well as basis (parameter) conversion operators, to be represented as banded matrices by exploiting the relations of the Jacobi polynomials -- they are extremely useful for solving a number of problems in not just 1D but 2D also, since one can utilise the sparsity of the relations that they hold to yield operator matrices that are efficient to solve \cite{doha2006efficient}.

Conversely, in a similar vein to the ultraspherical method for the unit interval \cite{olver2013fast, doha2009efficient}, a choice of OPs for the triangle are the \textit{Jacobi polynomials on the triangle}, which are a three-parameter family of polynomials in $(x,y)$ with differential operators acting in such a way so as to increment or decrement the parameter values \cite{dunkl2014orthogonal, koornwinder1975two, olver2019triangle}:
\bseqnnumber{
	P^{(a,b,c)}_{n,k}(x,y) := \tilde P^{(2k+b+c+1,a)}_{n-k}(x) \; (1-x)^k \; \tilde P^{(c,b)}_k\fpr({y \over 1-x}) \label{eqn:ds:triangleops}
}
for $(x,y) \in \bsset{(x,y) \in \R^2}{0 < x < 1, \: 0 < y < 1-x}$, where $\tilde P_k^{(a,b)}$ is the degree $k$ \textit{shifted Jacobi polynomial}\footnote{The shifted Jacobi polynomials are orthogonal on the interval $[0,1]$ and given by $\tilde P_k^{(a,b)}(x) = P_k^{(a,b)}(2x - 1)$ where $P_k^{(a,b)}$ is the standard degree $k$ Jacobi polynomial.} with $a,b,c$ being parameter values. Since there is no radial symmetry, differential operators for the triangle (as will be the case for the domains we investigate in this chapter) will be more complex than those for the disk.

@@

We can combine each system in \bsrefeqns{eqn:ds:jacobimatricesdefinitionx}{eqn:ds:jacobimatricesdefinitiony} into a block-tridiagonal system:
\bseqn{
	\renewcommand\arraystretch{1.3}
	&\begin{pmatrix}
		1 & & & \\
		B_0-G_0(x,y) & A_0 & & \\
		C_1 & B_1-G_1(x,y) & \quad A_1 \quad & \\
		& C_2 & B_2 - G_2(x,y)  & \ddots \\
		& & \ddots &\ddots
	\end{pmatrix}
	\bighdopabcd(x,y) \\
	&=
	\begin{pmatrix}
		\hdopabcd_{0,0} & 0 & 0 & 0 & \hdots  \\
	\end{pmatrix}^\top,
}
where we note $\hdopabcd_{0,0}(x,y) \equiv \genjac_0^{(a,b,c+d+1)} \: \jac_0^{(d,c)}$, and for each $n = 0,1,2\dots$,
\bseqn{
A_n &:= \begin{pmatrix}
		A^x_n \\
		A^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times(n+2)}, \quad
C_n := \begin{pmatrix}
		C^x_n \\
		C^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times n} \quad (n \ne 0), \nonumber \\
B_n &:= \begin{pmatrix}
		B^x_n \\
		B^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}, \quad
G_n(x,y) := \begin{pmatrix}
		xI_{n+1} \\
		yI_{n+1}
	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}.
}

For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e. such that $\Dnt A_n = I_{n+2}$. Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence for $n = 0,1,2,\dots$:
\bseqn{
	\begin{cases}
		\bighdopabcd_{-1}(x,y) := 0 \\
		\bighdopabcd_{0}(x,y) := \hdopabcd_{0,0} \\
		\bighdopabcd_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bighdopabcd_n(x,y) - \Dnt C_n  \,\bighdopabcd_{n-1}(x,y)
	\end{cases}.
}
Note that we can define an explicit $\Dnt$ as follows:
\bseqn{
\Dnt := \begin{pmatrix}
		\frac{1}{\alphaabcd_{n+1,0,1}} & &  \\
		& \ddots & & &  \\
		& & \ddots & & \\
		& & & \frac{1}{\alphaabcd_{n+1,n,1}} & \\
		\eta_{0} & 0 & \hdots & 0 & \eta_{1} & \hdots & \hdots & \eta_{n+1}
	    \end{pmatrix},
}
 where
 \bseqn{
	\eta_{n+1} &= \frac{1}{\betaabcd_{n,n,9}}, \\ 
	\eta_{n} &= -\frac{1}{\betaabcd_{n,n-1,9}} \big( \betaabcd_{n,n,8} \: \eta_{n+1} \big), \\
	\eta_j &= - \frac{1}{\betaabcd_{n,j-1,9}} \big( \betaabcd_{n,n+j+1,7} \: \eta_{j+2} + \betaabcd_{n,n+j,8} \: \eta_{j+1} \big) \qfor j = n-1,n-2,\dots,1, \\
	\eta_0 &= - \frac{1}{\alphaabcd_{n+1,0,1}} \big( \betaabcd_{n,1,7} \: \eta_{2} + \betaabcd_{n,0,8} \: \eta_{1} \big).
}
It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bighdopabcd_{0}(x,y)$  through $\bighdopabcd_{n}(x,y)$ in optimal $O(n^2)$ complexity.

@@

>>>>

For the disk, a review of spectral methods on the disk for solving the Poisson equation has been completed \cite{boyd2011comparing}, that looks at methods involving various bases for expanding functions in the disk including Zernike polynomials and Chebyshev polynomials. It was concluded that using Chebyshev polynomials yields mixed results, but that there was also no single advised basis to use, due to Zernike polynomials having a lack of radial fast transform at the time despite being found to yield accurate results. In the years since however, such radial fast transforms have been developed \cite{slevinsky2018use, sakai2009application, wilber2017computing}. The Zernike disk OPs can be written in terms of Jacobi polynomials \cite{boyd2011comparing}:
\bseqn{
	Z_n^m(r, \theta) := R_n^m(r) \sin(m\theta), \quad Z_n^{-m}(r, \theta) := R_n^m(r) \cos(m\theta),
}
where the radial part is
\bseqn{
	R_n^m(r) := r^m P^{(0,m)}_{(n-m)/2}(2r^2 - 1)
}
for $n - m$ even, while it is defined as $0$ for $n - m$ odd. A generalisation of the Zernike polynomials has been used for the development of a sparse spectral method on the disk \cite{vasil2016tensor}. These radial OPs are:
\bseqnnumber{
	Q_n^{k,m}(r) := C_n^{k,m} r^m P^{(k,m)}_n(2r^2 - 1), \label{eqn:ds:diskops}
}
where $C_n^{k,m}$ is some normalising constant so that they are orthonormal with respect to the inner product $\int_0^1 p(r) \; q(r) \; (1-r^2)^k \; r \; \D r$. $\{Q_n^{k,m}\}$ are therefore a two-parameter family of orthogonal polynomials on the unit disk. The classical Zernike polynomials are retrieved via setting $k = 0$. The use of polar coordinates here then allows for radially symmetric partial differential and other operators, as well as basis (parameter) conversion operators, to be represented as banded matrices by exploiting the relations of the Jacobi polynomials. Jacobi polynomials are extremely useful for solving a number of problems in both 1D and 2D, since one can utilise the sparsity of the relations that they hold to yield operator matrices that are efficient to solve \cite{doha2006efficient}.

In a similar vein to the ultraspherical method for the unit interval \cite{olver2013fast, doha2009efficient}, a choice of OPs for the triangle are the \textit{Jacobi polynomials on the triangle}, which are a three-parameter family of polynomials in $(x,y)$ with differential operators acting in such a way so as to increment or decrement the parameter values \cite{dunkl2014orthogonal, koornwinder1975two, olver2019triangle}:
\bseqnnumber{
	P^{(a,b,c)}_{n,k}(x,y) := \tilde P^{(2k+b+c+1,a)}_{n-k}(x) \; (1-x)^k \; \tilde P^{(c,b)}_k\fpr({y \over 1-x}) \label{eqn:ds:triangleops}
}
for $(x,y) \in \bsset{(x,y) \in \R^2}{0 < x < 1, \: 0 < y < 1-x}$, where $\tilde P_k^{(a,b)}$ is the degree $k$ \textit{shifted Jacobi polynomial}\footnote{The shifted Jacobi polynomials are orthogonal on the interval $[0,1]$ and given by $\tilde P_k^{(a,b)}(x) = P_k^{(a,b)}(2x - 1)$ where $P_k^{(a,b)}$ is the standard degree $k$ Jacobi polynomial.} with $a,b,c$ being parameter values. Since there is no radial symmetry, differential operators for the triangle (as will be the case for the domains we investigate in this chapter) will be more complex than those for the disk.

@@

We can combine each system in \bsrefeqns{eqn:ds:jacobimatricesdefinitionx}{eqn:ds:jacobimatricesdefinitiony} into a block-tridiagonal system:
\bseqn{
	\renewcommand\arraystretch{1.3}
	&\begin{pmatrix}
		1 & & & \\
		B_0-G_0(x,y) & A_0 & & \\
		C_1 & B_1-G_1(x,y) & \quad A_1 \quad & \\
		& C_2 & B_2 - G_2(x,y)  & \ddots \\
		& & \ddots &\ddots
	\end{pmatrix}
	\bighdopabcd(x,y) \\
	&=
	\begin{pmatrix}
		\hdopabcd_{0,0} & 0 & 0 & 0 & \hdots  \\
	\end{pmatrix}^\top,
}
where we note $\hdopabcd_{0,0}(x,y) \equiv \genjac_0^{(a,b,c+d+1)} \: \jac_0^{(d,c)}$, and for each $n = 0,1,2\dots$,
\bseqnnumber{
A_n &:= \begin{pmatrix}
		A^x_n \\
		A^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times(n+2)}, \quad
C_n := \begin{pmatrix}
		C^x_n \\
		C^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times n} \quad (n \ne 0), \label{eqn:ds:jacobisubmats1} \\
B_n &:= \begin{pmatrix}
		B^x_n \\
		B^y_n
	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}, \quad
G_n(x,y) := \begin{pmatrix}
		xI_{n+1} \\
		yI_{n+1}
	    \end{pmatrix} \in \R^{2(n+1)\times(n+1)}. \label{eqn:ds:jacobisubmats2}
}

For each $n = 0,1,2\dots$ let $\Dnt$ be any matrix that is a left inverse of $A_n$, i.e. such that $\Dnt A_n = I_{n+2}$. Multiplying our system by the preconditioner matrix that is given by the block diagonal matrix of the $\Dnt$'s, we obtain a lower triangular system \cite[p78]{dunkl2014orthogonal}, which can be expanded to obtain the recurrence for $n = 0,1,2,\dots$:
\bseqn{
	\begin{cases}
		\bighdopabcd_{-1}(x,y) := 0 \\
		\bighdopabcd_{0}(x,y) := \hdopabcd_{0,0} \\
		\bighdopabcd_{n+1}(x,y) = -\Dnt (B_n-G_n(x,y)) \bighdopabcd_n(x,y) - \Dnt C_n  \,\bighdopabcd_{n-1}(x,y)
	\end{cases}.
}
Note that we can define an explicit $\Dnt$ as follows:
\bseqnnumber{
\Dnt := \begin{pmatrix}
		\frac{1}{\alphaabcd_{n+1,0,1}} & &  \\
		& \ddots & & &  \\
		& & \ddots & & \\
		& & & \frac{1}{\alphaabcd_{n+1,n,1}} & \\
		\eta_{0} & 0 & \hdots & 0 & \eta_{1} & \hdots & \hdots & \eta_{n+1}
	    \end{pmatrix}, \label{eqn:ds:Dtdef}
}
 where
 \bseqn{
	\eta_{n+1} &= \frac{1}{\betaabcd_{n,n,9}}, \\ 
	\eta_{n} &= -\frac{1}{\betaabcd_{n,n-1,9}} \big( \betaabcd_{n,n,8} \: \eta_{n+1} \big), \\
	\eta_j &= - \frac{1}{\betaabcd_{n,j-1,9}} \big( \betaabcd_{n,n+j+1,7} \: \eta_{j+2} + \betaabcd_{n,n+j,8} \: \eta_{j+1} \big) \qfor j = n-1,n-2,\dots,1, \\
	\eta_0 &= - \frac{1}{\alphaabcd_{n+1,0,1}} \big( \betaabcd_{n,1,7} \: \eta_{2} + \betaabcd_{n,0,8} \: \eta_{1} \big).
}
It follows that we can apply $\Dnt$ in $O(n)$ complexity, and thereby calculate $\bighdopabcd_{0}(x,y)$  through $\bighdopabcd_{n}(x,y)$ in optimal $O(n^2)$ complexity.

@@

\begin{definition}\label{def:sc:clenshawmats}
	The matrices $-\Dnt (B_n-G_n(x,y,z))$, $\Dnt C_n$ defined via \bsrefeqns{eqn:ds:jacobisubmats1}{eqn:ds:Dtdef} are the recurrence coefficient matrices associated with the OPs $\{\hdopnkabcd\}$.
\end{definition}
